---
title: "Bellabeat Case Study - A Google Data Analytics Capstone Project Using Health Tracker Data"
author: "by Simona Cibulkova"
output: html_document
---

```{r ,echo=FALSE}
library(png)
library(grid)
img <- readPNG("Health_Tracker.png")
grid.raster(img)
```
  
## Background:

I have done this project as a part of the Google Data Analytics Professional Certificate. The goal of this study was to gain an insight into how consumers are using their smart devices and to make recommendations for the marketing strategy of the company.

**The Bellabeat company** is a high-tech manufacturer that designs innovative products that help women track their health and wellness. The company believes it has the potential to become a global leader in the smart device market, and analyzing smart device fitness data could create new growth opportunities for the company.

The company offers the following products: 
**the Bellabeat app** for tracking health data related to activity, sleep, stress, menstrual cycle, and mindfulness; 
**the Leaf** wellness tracker that can be worn as a bracelet, necklace, or clip, and syncs with the Bellabeat app to track activity, sleep, and stress levels; 
**the Time** watch that tracks activity, sleep, and stress levels and connects to the Bellabeat app for health insights; 
**the Spring** water bottle that connects to the Bellabeat app and tracks daily water intake to ensure people are adequately hydrated; the Bellabeat membership that provides 24/7 access to fully customized guidance on nutrition, activity, sleep, health, beauty, and mindfulness based on lifestyle and goals.  

#### Assignment:

As a guide to analyzing the data, I used the six steps of the data analysis process (ask, prepare, process, analyze, share, and act).  

#### Key Stakeholders:

* **Urška Sršen**: Co-founder and Chief Creative Officer of Bellabeat
* **Sando Mur**: Mathematician and co-founder of Bellabeat
  
#### Business Task:

Through the analysis of smart device usage data, we hope to gain insight into consumer use of non-Bellabeat smart devices. The Bellabeat marketing team would be given recommendations on how to use these trends or insights to craft new strategies.  
  
#### Business Objectives:

* What are the trends identified?
* How could these trends apply to Bellabeat customers?
* How could these trends help influence Bellabeat marketing strategy?  

  
#### Deliverables:

* A clear description of the business task.
* A description of all the data sources used.
* A description of any cleaning or manipulation done to the raw datasets.
* A summary of the results of the analysis.
* The supporting visualizations and key results.
  

## Dataset

For this case study, I used the data provided by Mobius that are available [online from Kaggle](https://www.kaggle.com/arashnic/fitbit). Using this data does not require a license as it is public domain.

These datasets were generated via Amazon Mechanical Turk by respondents to a distributed survey from 12th March to 12th May 2016. A total of 30 Health Tracker users agreed to submit personal tracker information, including activity intensity, a number of steps, heart rate, calories burned, and sleep monitoring. There are 18 .csv files containing data on the various tracking metrics broken down by days, hours, or minutes. Every data point has a timestamp and is assigned a user identifier.
  

#### Limitations of the Data

Data sources should be ROCCC, which stands for Reliable, Original, Comprehensive, Current, and Cited.

##### **Reliable** 
* After inspecting the data set, it was discovered that there were 33 unique user identifiers (not 30, as stated in the summary of the data set).
* A sample size of 30 (33) may not accurately reflect the product usage of the greater population and is not representative of the entire fitness population.
* Data belongs to Fitbit users whose consent was obtained to make their personal data available. Therefore, it is not a completely random sample. 
* Only data from the final month of the trial are included in the data set (April 12 - May 12).

##### **Original**
* The data was collected via a survey, i.e., by a third party, making it difficult to assess its originality.

##### **Comprehensive**
* Bellabeat tracked the actions of 33 participants over a period of 31 days, but not all participants tracked the same activities, so there are many gaps in the information. Only 24 participants tracked their sleep data, and only 8 participants tracked their weight. Additionally, there are no variables such as age and height included in the data that can be used to evaluate the effectiveness of activity tracking.
* It is not clear whether the distance is measured in miles or kilometers.

##### **Current**
* The data was collected in 2016. Due to this, users daily activities, fitness and sleeping habits, diet, and food consumption may have changed since then, and thus the data may be out of date.

##### **Cited**
* No available data.
  
Considering the small sample size and lack of demographic data, this case study is more of a training exercise than a legitimate study of actionable insights due to the high margin of error.  

Consider this scenario: your company wants to market only to Czechs. The Czech Republic is the only EU country where more than a third of people aged 16 to 74 use smartwatches and other similar devices. That is the equivalent of 2.8 million people. This is supported by data published by [Eurostat last year](https://ec.europa.eu/eurostat/web/products-eurostat-news/-/ddn-20210225-1?redirect=%2Feurostat%2F). We would need a sample size of 385 to be able to perform an accurate analysis (with a 95% confidence level and a 5% margin of error) of Czech fitness tracker user behavior. [Link to Sample Size Calculator](https://www.surveymonkey.com/mp/sample-size-calculator). This analysis uses data from 33 users, giving us a margin of error of 18% for an analysis of the public, which is higher than the general margin of error range of 4% - 8%. [Link to Margin of Error Calculator](https://www.pollfish.com/margin-of-error-calculator).

While we may have enough samples for a smaller demographic, such as a smaller area of population, we are unable to accurately tie the data to a particular region/location since we do not have any demographic information.  
  

## Setting up an environment

In R, the required packages have been installed and uploaded.

```{r }
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("tidyr")
# install.packages("readr")
# install.packages("lubridate")
# install.packages("ggplot2")
# install.packages("tinytex")
# install.packages("ggpubr")
# install.packages("magrittr")
# install.packages("ggthemr")
# install.packages("psych")
# install.packages("gridExtra")
# install.packages("png")
# install.packages("grid")
# install.packages("hrbrthemes")
```

```{r loadlib, echo=T, results='hide', message=F, warning=F}
library(tidyverse)
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(tinytex)
library(ggpubr)
library(magrittr)
library(ggthemr)
library(psych)
library(gridExtra)
library(hrbrthemes)
```
  

## Data selection

For further analysis, the following datasets were selected.  

```{r echo=T, results='hide', message=F, warning=F}

daily_activity <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/dailyActivity_merged.csv')

hourly_steps <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/hourlySteps_merged.csv')

daily_sleep <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/sleepDay_merged.csv')

hourly_calories <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/hourlyCalories_merged.csv')

hourly_intensity <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/hourlyIntensities_merged.csv')

weight <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/weightLogInfo_merged.csv')

heartrate <- read.csv('C:/Users/scibu/OneDrive/Plocha/Bellabeat/data/heartrate_seconds_merged.csv')
```
  

## Data cleansing and transformation

The data were processed by cleaning and ensuring they are correct, relevant, complete, free of errors, and not displaying any outliers by:

* exploring and observing the data,
* identifying missing values, processing them, and analyzing them,
* formatting data - converting it into a specific format,
* conducting preliminary statistical analyses.  

  
## Examining the data and becoming familiar with it

To see the structure of the data, the data frames were initially displayed as strings, tibbles or glimpses.  
```{r}
#str(hourly_steps)
str(daily_activity)
#str(daily_sleep)
#str(hourly_calories)
#str(hourly_intensity)
#str(weight)
#str(heartrate)
```

```{r}
#as_tibble(hourly_steps)
#as_tibble(daily_activity)
as_tibble(daily_sleep)
#as_tibble(hourly_calories)
#as_tibble(hourly_intensity)
#as_tibble(weight)
#as_tibble(heartrate)
```

```{r}
#glimpse(hourly_steps)
#glimpse(daily_activity)
#glimpse(daily_sleep)
#glimpse(hourly_calories)
#glimpse(hourly_intensity)
glimpse(weight)
#glimpse(heartrate)
```
  

### Counting the number of users

Number of unique users in each dataset was counted.  

```{r}
n_distinct(daily_activity$Id)
n_distinct(hourly_calories$Id)
n_distinct(hourly_intensity$Id)
n_distinct(hourly_steps$Id)
n_distinct(daily_sleep$Id)
n_distinct(weight$Id)
n_distinct(heartrate$Id)
```
  

##### Observations:

* The following Daily and Hourly Activity data is available for the period of 2016-04-12 to 2016-05-12 with 33 unique IDs.
* Although the dataset should have included 30 respondents, there are 33 entries. It is possible that some respondents created additional identifiers while filling in the survey. 
* There are just 24 unique IDs in daily sleep data from 2016-04-12 to 2016-05-12.
* With 8 unique IDs, weight data is available from 2016-04-12 to 2016-05-12.
* With 14 unique identifiers, data on heart rate variability is available for the period from 2016-04-12 to 2016-05-12.
* In other words, 73 percent of users are willing to track sleep time, only 24 percent are willing to track weight, and only 42% wish to track heart rate variability.
* For selected features, users are checking at most their daily activity, hourly calories, hourly intensity, and hourly steps.  
  

### Checking for NULL values

```{r}
sum(is.na(daily_activity),is.na(hourly_calories),is.na(hourly_intensity),is.na(hourly_steps),is.na(daily_sleep),is.na(heartrate))
sum(is.na(weight))

which(is.na(weight))
```
  

##### Observations:

Only the weight dataset contains null values. Identifying which column has null values:

```{r}
colnames(weight)[colSums(is.na(weight)) > 0]
weight <- select(weight, -Fat)
```

The only column that contained null values was Fat column, which was then removed.  
  
### Creating a Venn diagram 

```{r echo=T, results='hide', message=F, warning=F}
daily_ids <- unique(daily_activity$Id)
hourly_ids <- unique(hourly_calories$Id)
sleep_ids <- unique(daily_sleep$Id)
weight_ids <- unique(weight$Id)
heartrate_ids <- unique(heartrate$Id)


library(VennDiagram)
venn <- venn.diagram(x = list("Steps" = daily_ids,
                              "Sleep" = sleep_ids,
                              "Weight" = weight_ids,
                              "Heartrate" = heartrate_ids),
                     filename = "features_venn_diagram.png",
                     imagetype = "png",
                     fill = c("blue", "green", "yellow", "red"),
                     fontfamily = "sans",
                     fontface = "bold",
                     cat.fontface = "bold",
                     cat.fontfamily = "sans",
                     lwd = 3,
                     cex = 2
                     )
```


```{r ,echo=FALSE}
img <- readPNG("features_venn_diagram.png")
 grid.raster(img)
```
  

##### Observations:

Tracking steps is an essential feature, used by 100 percent of users. With 72.7 percent of users, sleep tracking is the second most popular feature. Less than half (42,4 percent) of users use heart rate tracking. Fewer than one-quarter (24.2 percent) of users track weight. Just 9% of users use all four tracking features simultaneously.

The following analyses will focus on tracking steps and sleep as these are the most popular features.  
  

### Data manipulation and transformation

Some data frames had to be formatted in YYYY-MM-DD so R could work well with them. The date columns data types were then changed from character to date time.

Activity Date is the date column but was found to be formatted as a character type. It should be formatted as a data type for better analysis.

A new column for a time of day was created for the hourly steps data frame in order to analyze the data by time of day (i.e. 1:00 AM, 2:00 AM, etc.).

```{r echo=T, results='hide', message=F, warning=F}

# formating original date+time column with am/pm
daily_sleep$Date <- parse_date_time(daily_sleep$SleepDay, '%m/%d/%Y %I:%M:%S %p')

# formating date from character type to date type
daily_activity$Date <- parse_date_time(daily_activity$ActivityDate, '%m/%d/%Y')

# formating date + time (in am/pm)
hourly_intensity$ActivityHour <- parse_date_time(hourly_intensity$ActivityHour, '%m/%d/%Y %I:%M:%S %p')
hourly_intensity$Date <- format(as.POSIXlt(hourly_intensity$ActivityHour), "%m/%d/%Y")
hourly_intensity$Time <- format(as.POSIXlt(hourly_intensity$ActivityHour), "%H:%M:%S")
hourly_intensity$Date <- parse_date_time(hourly_intensity$Date, '%m/%d/%Y')

# separating date+time, deleting time (which is irrelevant for the analysis) and formating date
weight <- weight %>% separate(Date, c("Date", "Time"), " ")
weight <- subset(weight, select=-Time)
weight$Date <- parse_date_time(weight$Date, '%m/%d/%Y')
```
  

For each observation, a column representing the day of the week (e.g. Monday, Tuesday, etc.) was added.   

```{r}
daily_activity$Weekday <- format(as.Date(daily_activity$Date), "%A")
daily_activity$Weekday <- recode(daily_activity$Weekday, "pondělí" = "Monday","úterý" = 'Tuesday', "středa" = 'Wednesday', "čtvrtek" = "Thursday", "pátek" = "Friday", "sobota" = "Saturday", "neděle" = "Sunday")
daily_activity$Weekday <- ordered(daily_activity$Weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

daily_sleep$Weekday <- format(as.Date(daily_sleep$Date), "%A")
daily_sleep$Weekday <- recode(daily_sleep$Weekday, "pondělí" = "Monday","úterý" = 'Tuesday', "středa" = 'Wednesday', "čtvrtek" = "Thursday", "pátek" = "Friday", "sobota" = "Saturday", "neděle" = "Sunday")
daily_sleep$Weekday <- ordered(daily_sleep$Weekday, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
```
  

To record the total daily activity, a column was created for the total minutes.  

```{r}
daily_activity$TotalMinutes <- rowSums(daily_activity [,11:14])
daily_activity$TotalActivityMinutes <- rowSums(daily_activity [,11:13])
```
  

Estimation of the amount of time the user spends asleep in bed:  

```{r}
daily_sleep <- mutate(daily_sleep,PercentOfSleep = (TotalMinutesAsleep/TotalTimeInBed)*100)
```
  

### Initial table summaries:

#### Basic statistics includes:
* mean (average)
* std (standard deviation)
* min and max
* 25, 50 and 75 percentiles 
  

#### *Weight summary*

```{r}
weight %>%
  select(WeightKg,BMI) %>%
  summary()
```
  

##### Observations:

The weight data of only eight participants has been logged. The average daily weight log stands at 72.04 kg, and the average body mass index stands at 25.19, which is at or near the bottom of the overweight range of 25 to 29.9.  
  

#### *Daily activity summary*

```{r}
# explore total activity
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes, Calories) %>%
  summary()

# explore numbers of activity intensity
daily_activity %>%
  select(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes) %>%
  summary()
```
  

##### Observations:

The summary shows that the average user is taking 7,638 steps a day, which is less than the recommended 10,000 steps. According to the 3rd quartile of TotalSteps stats, 75% of users maintain 10,000 steps per day. Users get 21.16 minutes of very active or vigorous activity on average per day, which translates into 148.12 minutes per week. The average user meets the recommended 75 minutes of vigorous activity per week. In order to counteract the effects of sitting up to 10 hours a day, scientists recommend performing moderate to vigorous activity for at least 40 minutes each day. In this study, participants spend an average of 991.2 minutes sedentary, which equals 16.52 hours a day (including roughly 7 hours of sleep). Such long periods of inactivity can cause health problems. According to this summary, the average individual burns 2304 calories a day. It takes 3,500 calories to lose one pound of fat, so users intending to lose weight should burn more than this.  
  

#### *Average activities during different weekdays*

```{r}
# merge datasets
daily_sleep <- merge(daily_activity, daily_sleep, by = c("Id", "Date", "Weekday"))

# check for number of unique ids (should be 24)
paste("Unique ids:", n_distinct(daily_sleep$Id))
```


```{r}
avg_daily_sleep_and_activity <- daily_sleep %>% 
  group_by(Weekday) %>% 
  summarize(avg_calories = mean(Calories),
            avg_steps = mean(TotalSteps),
            avg_distance = mean(TotalDistance),
            avg_running_min = mean(VeryActiveMinutes),
            avg_jogging_min = mean(FairlyActiveMinutes),
            avg_walking_min = mean(LightlyActiveMinutes),
            avg_sedentary_min = mean(SedentaryMinutes),
            avg_asleep_min = mean(TotalMinutesAsleep),
            avg_in_bed_min = mean(TotalTimeInBed),
            avg_awake_in_bed_min = mean(TotalTimeInBed-TotalMinutesAsleep),
            avg_sleep_sessions = mean(TotalSleepRecords))

as_tibble(avg_daily_sleep_and_activity)
```
  

##### Observations:

The above table shows that users are walking (light activity) at least three hours a day, every day, during the week. A relatively high number of minutes are spent sedentary each day. 
  
#### *Sleep record summary*

```{r}
# explore total sleep numbers
daily_sleep %>%
  select(TotalSleepRecords, TotalMinutesAsleep, TotalTimeInBed, PercentOfSleep) %>%
  summary()
```


##### Observations:

The average amount of time spent in bed is 458.6 minutes or 7.6 hours.  

```{r}
# Correlation plot with steps and sleep data
daily_sleep %>% 
  select(c(Calories, TotalSteps, TotalDistance, TotalMinutesAsleep, TotalTimeInBed, SedentaryMinutes, LightlyActiveMinutes, VeryActiveMinutes, FairlyActiveMinutes)) %>% 
  rename(Steps = TotalSteps, Km = TotalDistance, Sleep = TotalMinutesAsleep, Laying = TotalTimeInBed, Sit = SedentaryMinutes, Walking = LightlyActiveMinutes, Run = VeryActiveMinutes, Jogging = FairlyActiveMinutes) %>% 
  corPlot(cex = 1.1, show.legend = FALSE)
```
  

##### Observations

There is a notable correlation between calories burned and time spent running (0.61), whereas jogging (0.18) and walking (0.12), however, have a much smaller correlation. As a result, running is a more effective calorie-burning activity than walking or jogging.

There is a moderate negative correlation between sitting time and walking time (-0.26). It has been found that the more a person sleeps, the less time they spend sitting.

In agreement with expectations, distance and number of steps are highly correlated (0.98). There are significant positive correlations between these variables and calories burned, time spent walking, time spent jogging, and time spent running. The amount of time spent sleeping seems to have little influence (-0.19) on how many steps a person takes.  
  

#### *Visualisation based on users activity*

```{r}
group_by_usertype <- 
    daily_activity %>%
    summarise(
        user_type = factor(case_when(
        SedentaryMinutes > mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Sedentary",
        SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes > mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Lightly Active",
        SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes > mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Fairly Active",
        SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes > mean(VeryActiveMinutes) ~ "Very Active",
    ),levels=c("Sedentary", "Lightly Active", "Fairly Active", "Very Active")), Calories, Id=Id) %>%
    drop_na()

# plotting user categories
options(repr.plot.width = 10, repr.plot.height = 8)
ggthemr("light")
group_by_usertype %>%
group_by(user_type) %>%
summarise(total = n()) %>%
mutate(totals = sum(total)) %>%
group_by(user_type) %>%
summarise(total_percent = total / totals) %>%
ggplot(aes(user_type,y=total_percent, fill=user_type)) +
    geom_col()+
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position="none") +
    labs(title="Users' Activity Distribution", y = "Percentage", x = "Type of Activity", caption = paste0("Data from 33 users from 4/12/2016 to 5/12/2016"))
```
  

##### Observations:

The graph illustrates that users spend most of their day in a sedentary position. This is followed by light activities such as walking. Movements that are very active (running) and fairly active (jogging) are below 10%.    

#### *Box plot of calories burned by user type*

```{r}
options(repr.plot.width = 10, repr.plot.height = 8)
ggthemr("dust")
ggplot(group_by_usertype, aes(user_type, Calories, fill=user_type)) +
    geom_boxplot() +
    theme(legend.position="none") +
    labs(title="Calories Burned by User Type", x="User Type")
```
  

##### Observations:

Exercise intensity and calories burned are positively correlated, as can be seen in the graph. Very active people (those who run) burn more calories than moderately active ones (those who jog), lightly active ones (those who walk), and sedentary ones (those who sit).    

#### *Visualizing number of steps a day*

```{r}

daily_activity_steps_grouped <- daily_activity %>%
  group_by(Weekday) %>%
  summarise(mean_total_steps = mean(TotalSteps))

head(daily_activity_steps_grouped)

options(repr.plot.width = 10, repr.plot.height = 8)
ggthemr("grass")
ggplot(daily_activity_steps_grouped, aes(x = Weekday, y = mean_total_steps, fill = mean_total_steps)) +
  geom_col() + 
  theme_bw()+
  labs(y = "Average Steps", 
      x = "Weekday", 
      title = "Average steps by week",
      fill = "TotalSteps")
```
  

##### Observations:

On Tuesday and Saturday, users were most active, and on Sunday they were least active.    

#### *Plotting hourly intensity to see what time of day users were most active*

```{r echo=T, message=F, warning=F}
hour_mean <- hourly_intensity %>% 
  summarize(avg_total_intensity = mean(TotalIntensity))
hour_mean

options(repr.plot.width = 10, repr.plot.height = 8)
ggthemr('flat')
hourly_intensity %>%
  group_by(Time) %>%
  summarise(average_total_intensity = mean(TotalIntensity)) %>%
    ggplot(aes(Time, y=average_total_intensity, fill=average_total_intensity)) + 
        geom_histogram(stat = "identity") +
        theme_bw()+
        theme(axis.text.x = element_text(angle = 45)) +
        labs(title="Average Total Intensity During the Day",
            x="Time of the Day",
            y="Average Intensity")+
        annotate("text", x = 4, y = 13, label = "Mean hourly intensity: 12", color = 'black', size = 3)+
        geom_hline(yintercept = mean(hour_mean$avg_total_intensity),color="brown")
```
  

##### Observations:

There was a peak in activity around 7 PM with the most activity between 5 PM and 8 PM.    

#### *Visualizing Calories and Total Steps*

Examining the relationship between Steps and calories burned.  

```{r}
ggthemr('lilac')
ggplot(data=daily_activity, aes(x=TotalSteps, y=Calories))+
  geom_point()+
  geom_smooth(method=lm)+
  labs(title = "Total Steps and Calories Burned", x="Total Steps", y="Calories")+
  scale_x_comma()
```
  

##### Observations:

The plot shows a positive correlation between total steps and calories burned. There do, however, seem to be some outliers that do not appear to follow the correlation. These include observations of 0 steps with zero to minimal calories burned and an observation of more than 35,000 steps with fewer than 3,000 calories burned.    

#### *Checking for a correlation between the different activities and the calories burnt*

```{r}
ggthemr('light')
plot2 <- ggplot(data=daily_activity, aes(x=FairlyActiveMinutes, y=Calories))+
  geom_point()+
  geom_smooth()+
  labs(title="Calories Burned in Fairly Active Minutes", x= "Fairly Active Minutes")
plot1 <- ggplot(data=daily_activity, aes(x=LightlyActiveMinutes, y=Calories))+
  geom_point()+
  geom_smooth()+
  labs(title="Calories Burned in Lightly Active Minutes", x= "Lightly Active Minutes")

grid.arrange(arrangeGrob(plot1, plot2, ncol = 2),nrow = 1)
```

```{r}
ggthemr('light')
ggplot(data=daily_activity, aes(x=VeryActiveMinutes, y=Calories))+
  geom_point()+
  geom_smooth()+
  labs(title="Calories Burned in Very Active Minutes", x= "Very Active Minutes")
```
  

##### Observations

The intensity of exercise (walking, running, jogging) is associated with the number of calories burned. Very active movement is the best predictor of calories burned.   

#### *Sleeping time during the week*  

```{r}
options(repr.plot.width = 8, repr.plot.height = 8)
ggthemr("light")
ggplot(data = avg_daily_sleep_and_activity)+ geom_col(mapping = aes(x = Weekday, y = avg_asleep_min, fill = avg_asleep_min))+
  labs(title = "Average Sleep eachDring the Week", 
       x="Week day", y = "Average Sleep Minutes", 
       fill = "Total Sleep Minutes")
```
  

##### Observations:

On weekends, especially on Sunday, people tend to sleep more. During working days, an average of 400 minutes (6-7 hours) is maintained.    

#### *Total time in bed and total time of sleeping*

```{r}
options(repr.plot.width = 8, repr.plot.height = 8)
ggthemr("flat")
ggplot(daily_sleep, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + 
  geom_point()+
  geom_smooth()+
  labs(title="Time in Bed vs. Minutes Asleep", x="Total Time of Sleep", y="Total Time Spent in Bed")
```
  

##### Observations:

Time asleep is positively related to total time spent in bed, but some points are outliers, indicating that some people have trouble falling asleep even after spending a long time in bed.    


### Insights:

* Users were most active on Saturdays and Tuesdays, and least active on Sundays. Over the course of the week, they were most active between 5 PM and 8 PM, with a peak around 7 PM (probably when and after getting home from work).  

* Calorie burnt rate varies with the type of exercise. Very active movement and distance traveled are the best predictors of calories burned. A positive correlation exists also between total steps and calories burned.

* On average, users burned 2,303 calories and walked 7,637 steps per day. In terms of health benefits, that is almost 2400 steps fewer than the recommended 10,000 steps per day.

* The average number of sedentary hours per day was 16.5 hours, including an average of 7 hours of sleep per day.

* The majority of users kept track of sleep time - 73% did so.

* The average body mass index recorded was 25.19, which is in the overweight category of 25 to 29.9. It suggests that most people are either not following a healthy diet or not exercising enough. However, only 24% of participants had logged their weight. The reasons for this need to be explored further. Are the weight logging procedures complicated? Are the participants not interested in tracking their weight progress?

* The heart rate variability was tracked by only 42% of participants. A more thorough investigation is needed to discover why. Why aren not participants tracking their weight progress?
  

### Recommendations:

* To conduct further analysis, more data should be collected. There is no demographic information in the current data. Hence, a larger sample size would be beneficial for conducting further analysis.

* Educating and empowering users about fitness benefits (like a type of movement and calories burned) via the Bellabeat app could encourage them to exercise more. In addition, the alert notifications can motivate the user to reach the daily goal of 10,000 steps.

* Bellabeat can also provide notifications to encourage users to exercise on weekends or in the evening. I would suggest sending users a variety of reminders or strategies like suggesting they spread out their recommended 30 minutes of activity throughout the day and week and encouraging them to do some exercise if they have been sedentary for too long (for example, taking a walk every 2-3 hours for 5-10 minutes or getting up and moving every 30 minutes to avoid risks associated with prolonged sitting and to boost productivity). The individual could be advised to perform 75 minutes of vigorous activity a week and to receive alert notifications.

* Bellabeat could offer more than just steps, sleep time, weight, and heart rate tracking. It could offer sessions on eating habits, different types of exercises, etc.

* The notifications can help users stay on track in their efforts to burn the necessary calories and reach their weight loss goals if they are trying to lose weight.

* Bellabeat should make it easier to record weight and BMI. An example would be a wireless connection between a weighing machine and an app, or a built-in camera that allows customers to take a picture of their weight on a digital scale and enter it automatically into the log. Customers may be encouraged to weigh themselves by having a reminder icon or a beeping sound go off at selected times throughout the day.

* Similarly to logging weight, new methods should be developed to make tracking sleep easier for customers. It is also important to emphasize the relationship between sleep quality and quantity and health and beauty.

* For app users who want to improve their sleep time, Bellabeat could send a notification via the app prompting them to go to bed at a specific time.

* Exercise with a friend is a great way to feel more motivated, to try new types of exercises, and to stay consistent. Consequently, social media workout apps could help users reach their goals by connecting them with friends and other users. By creating a supportive women's group ready to prioritize their health, the Bellabeat app could become women's go-to workout app. In such an environment, users could also post their favorite workouts, wellness tips, healthy recipes, etc. |||||| Users could add friends and view each other’s activity too. They could follow weekly fitness and wellness challenges.

* Heartrate was not studied here so I would just recommend enabling alert notifications if their resting heart rate varied significantly from their normal level.